---
title: "urchinDataDiscretization"
author: "Jorge Arroyo-Esquivel"
date: "7/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Let's install tidyverse to analyze the data
```{r}
library(tidyverse)
library(here)
```

### Let's load the data
```{r}
data <- read.csv(paste(here(),'Raw Data/Urchin Data/Reef Check Urchin Data.csv',sep='/'))
summary(data)
```

We only care about some Latitudes (between 38.5 and 39.75)

### Let's clean and summarize the data that we want
```{r}
data$Lat <- round(data$Lat,digits=4)
cleanData <- data %>% filter(between(Lat,38.5,39.75)) %>% group_by(Year,Lat) %>% summarise(TotalObservations=n(),MeanCount=mean(as.numeric(Count)))
cleanData %>% filter(Year==2007) %>% ggplot()+geom_point(aes(Lat,MeanCount))
```

Data is pretty scarce.

### Let's repeat the process done on kelp of interpolating data as a first approximation
```{r}
stepSize <- 1e-4
n <- ceiling((39.75-38.5)/stepSize)
x <- round(seq(38.5,39.75,length.out = n),digits=4)
years <- unique(cleanData$Year)

U <- NULL
for(i in years){
  currYear <- cleanData %>% filter(Year==i) %>% ungroup() %>%  select(Lat,MeanCount)
  Ut <- data.frame(Lat=x) %>% left_join(currYear)
  for(j in 1:n){
    if(is.na(Ut$MeanCount[j])){
      gap <- Ut$MeanCount[j:min(j+60,n)]
      
      Ut$MeanCount[j] <- ifelse(is.nan(mean(gap,na.rm = TRUE)),0,mean(gap,na.rm = TRUE))
    }
  }
  U <- cbind(U,Ut$MeanCount)
}
```

### Let's plot this and see how it looks
```{r}
plot(x,U[,10],type='l')
```

This suggests we should work on a more local scale, where urchin data is easily interpolabable and Ut(x) is reliable

### Let's save these results
```{r}
save(x,U,file='Utx.rdata')
```

